---
title: "Statistical Analysis and Modeling - Coursework 1"
output:
  html_document:
    df_print: paged
date: "01/12/2022"
owner: Umer Bin Usman
---

## Installing Important Libraries


```{r}
install.packages("ggplot2")
install.packages("gridExtra")
install.packages("moments")
install.packages("devtools")
install.packages("corrplot")
install.packages("dplyr")
install.packages("randomForest")
install.packages("Metrics")
install.packages("car")
install.packages("performance")
```


## Loading Libraries
```{r}
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)
library(scales)
library(moments)
library(devtools)
library(corrplot)
library(reshape2)
library(dplyr)
library(caret)
library(car)
library(randomForest)
library(Metrics)
library(performance)
```

## Uploading my Dataset

```{r}
main_dataset <- read.csv("D:/Data Science/Statistical Analysis and Modeling/Coursework 1/House Price/Dataset/Main Dataset.csv", stringsAsFactors = FALSE)
```

## Printing the first 10 instances of my dataset from each attribute

```{r}
View(main_dataset)
head(main_dataset, 10)
```

## Filtering Unwanted Attributes

```{r}
f_data <- dataset[, c("Id", "MSSubClass", "LotArea", "LandContour", "Utilities", "Neighborhood", 
                             "Condition1", "Condition2", "BldgType", "OverallQual", "OverallCond", "YearBuilt",
                             "YearRemodAdd", "Heating", "HeatingQC", "CentralAir", "Electrical", "KitchenQual",
                             "GarageType", "GarageCars", "MoSold", "YrSold", "SalePrice")]
```

## Printing Filtered Attributes

```{r}
f_data
```

## Checking the filtered attributes
```{r}
names(f_data)
```
## Changing categorical data from string to factor
```{r}
f_data$LandContour <- as.factor(f_data$LandContour)
f_data$Utilities <- as.factor(f_data$Utilities)
f_data$Neighborhood <- as.factor(f_data$Neighborhood)
f_data$Condition1 <- as.factor(f_data$Condition1)
f_data$Condition2 <- as.factor(f_data$Condition2)
f_data$BldgType <- as.factor(f_data$BldgType)
f_data$Heating <- as.factor(f_data$Heating)
f_data$HeatingQC <- as.factor(f_data$HeatingQC)
f_data$CentralAir <- as.factor(f_data$CentralAir)
f_data$Electrical <- as.factor(f_data$Electrical)
f_data$KitchenQual <- as.factor(f_data$KitchenQual)
f_data$GarageType <- as.factor(f_data$GarageType)
```


## Checking the empty values Dataset

```{r}
is.na(f_data)
```

## Finding the sum of empty values in whole dataset

```{r}
sum(is.na(f_data))
```

## Removing the rows with empty values in whole dataset

```{r}
f_data1 <- na.omit(f_data)
```

## Finding the sum of empty values in whole dataset

```{r}
sum(is.na(f_data1))
```

## Comparing the sum of empty values

```{r}
sum(is.na(f_data))
sum(is.na(f_data1))
```

## Checking the structure of filtered dataset

```{r}
str(f_data1)
```

## Checking the summary of filtered dataset

```{r}
summary(f_data1)
```

## Creating bar plot of all the categorical attributes
```{r}
data <- data.frame(f_data1)
chart1 = ggplot(data, aes(x=LandContour)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart2 = ggplot(data, aes(x=f_data1$Utilities)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart3 = ggplot(data, aes(x=f_data1$Neighborhood)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart4 = ggplot(data, aes(x=f_data1$Condition1)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart5 = ggplot(data, aes(x=f_data1$Condition2)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart6 = ggplot(data, aes(x=f_data1$BldgType)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart7 = ggplot(data, aes(x=f_data1$Heating)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart8 = ggplot(data, aes(x=f_data1$HeatingQC)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart9 = ggplot(data, aes(x=f_data1$CentralAir)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart10 = ggplot(data, aes(x=f_data1$Electrical)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart11 = ggplot(data, aes(x=f_data1$KitchenQual)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
chart12 = ggplot(data, aes(x=f_data1$GarageType)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust =1))
```

## Showing all the barplots
```{r}
chart1
chart2
chart3
chart4
chart5
chart6
chart7
chart8
chart9
chart10
chart11
chart12
#grid.arrange(chart1, chart2, chart3, 
 #            ncol=2, nrow=3)
```
## Creating the Boxplot to check the prices in different area
```{r}
#chart1 = ggplot(data, aes(x= Neighborhood, y= SalePrice)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust =1))
chartX = ggplot(data, aes(x= Neighborhood, y= SalePrice)) + geom_boxplot() + geom_hline(aes(yintercept=mean(data$SalePrice)), colour = "red", linetype= "dashed", lwd=0.5) + scale_y_continuous(labels = dollar_format()) + theme(axis.text.x = element_text(angle = 90, hjust =1))
chartX
```

## Creating the density graph to check the distribution of data and skewness
```{r}
chart_d1 = ggplot(data = data, aes(x = MSSubClass)) + geom_density( fill = "yellow", color = "black")
chart_d2 = ggplot(data = data, aes(x = LotArea)) + geom_density( fill = "yellow", color = "black")
chart_d3 = ggplot(data = data, aes(x = OverallQual)) + geom_density( fill = "yellow", color = "black")
chart_d4 = ggplot(data = data, aes(x = OverallCond)) + geom_density( fill = "yellow", color = "black")
chart_d5 = ggplot(data = data, aes(x = YearBuilt)) + geom_density( fill = "yellow", color = "black")
chart_d6 = ggplot(data = data, aes(x = YearRemodAdd)) + geom_density( fill = "yellow", color = "black")
chart_d7 = ggplot(data = data, aes(x = GarageCars)) + geom_density( fill = "yellow", color = "black")
chart_d8 = ggplot(data = data, aes(x = MoSold)) + geom_density( fill = "yellow", color = "black")
chart_d9 = ggplot(data = data, aes(x = YrSold)) + geom_density( fill = "yellow", color = "black")
```


```{r}
chart_d1
chart_d2
chart_d3
chart_d4
chart_d5
chart_d6
chart_d7
chart_d8
chart_d9
```
## Printing all the skewness values
```{r}
print(paste0("MSSubClass = " , skewness(data$MSSubClass)))
print(paste0("LotArea = " , skewness(data$LotArea)))
print(paste0("OverallQual = " , skewness(data$OverallQual)))
print(paste0("OverallCond = " , skewness(data$OverallCond)))
print(paste0("YearBuilt = " , skewness(data$YearBuilt)))
print(paste0("YearRemodAdd = " , skewness(data$YearRemodAdd)))
print(paste0("GarageCars = " , skewness(data$GarageCars)))
print(paste0("MoSold = " , skewness(data$MoSold)))
print(paste0("YrSold = " , skewness(data$YrSold)))
```
## Correlation  
```{r}
num_data <- select_if(data, is.numeric)
corr_mat <- round(cor(num_data), 2)

melted_corr_mat <- melt(corr_mat)
head(melted_corr_mat)

ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) + theme(axis.text.x = element_text(angle = 90, hjust =1))
```

## Looking for Outliers
I will make scatter plot between yearbuilt and garaegcars

## Modeling and Prediction

## Structure of Filtered Data
```{r}
str(f_data1)
```
## Finding missing values
```{r}
sum(!complete.cases(f_data1))
```
## Splitting Dataset
```{r}
data_train <- f_data1[1:1300,]
data_train

data_test <- f_data1[1301:2757, ]
#data_test <- subset (data_test, select = -SalePrice)
data_test
```
## Primary Model - Random Forest
```{r}
set.seed(1)

ranfor_model <- randomForest(x = data_train[, 1:22], 
                          y = data_train$SalePrice,
                          ntree = 100)

ranfor_model
```
## Ploting Error and Variable importance chart
```{r}
plot(ranfor_model)
varImpPlot(ranfor_model)

```
## Applying Model on test data
```{r}
ranfor_pred <- predict(ranfor_model, data_test)
View(ranfor_pred)
```

## Cross Validation of Random Forest
```{r}

print(paste0("Root Mean Square Error (RMSE) = ", RMSE(ranfor_pred, data_test$SalePrice)))
print(paste0("R Square Error (R2) = ", R2(ranfor_pred, data_test$SalePrice)))
print(paste0("Mean Absolute Error (MAE) = ", MAE(ranfor_pred, data_test$SalePrice)))
```


## Second Model - Multiple Linear Regression
```{r}
set.seed(1)

mul_reg <- lm(data_train$SalePrice ~., data = data_train)
summary(mul_reg)
#plot(mul_reg)

```

## Plotting Quantile Plot
```{r}
plot(mul_reg)
```

## Applying model on test data
```{r}
mulreg_pred <- predict(mul_reg, data_test)
View(mulreg_pred)
```

## Cross Validation of Multiple Linear Regression 
```{r}
print(paste0("Root Mean Square Error (RMSE) = ", RMSE(mulreg_pred, data_test$SalePrice)))
print(paste0("R Square Error (R2) = ", R2(mulreg_pred, data_test$SalePrice)))
print(paste0("Mean Absolute Error (MAE) = ", MAE(mulreg_pred, data_test$SalePrice)))
```

























